---
title: "P8105_hw2_yx2507"
output: github_document
---
Yuqing Xue
yx2507 Oct.2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1  
## 1.1 Read and clean the Mr. Trash Wheel sheet:
* specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
* use reasonable variable names
* omit rows that do not include dumpster-specific data
* round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r，problem1 trash wheel data, message = FALSE}
trash_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx") %>%
  janitor::clean_names() %>%
  filter(dumpster != "NA") %>%
mutate(sports_balls = as.integer(sports_balls))

## check
view(trash_data)

```
## 1.2 Read and clean precipitation data for 2017 and 2018. For each, omit rows without precipitation data and add a variable year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).
```{r, problem1 precipitation , message = FALSE}

data_rain_2018 = 
    read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = '2018 Precipitation', range = "A2:B15") %>% 
    janitor::clean_names() %>% 
    filter(total != "NA") %>% 
    mutate(year = 2018)

data_rain_2017=
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = '2017 Precipitation', range = "A2:B15") %>% 
    janitor::clean_names() %>% 
    filter(total != "NA") %>% 
    mutate(year = 2017)

```

## combine and get a`month.name` variable
```{r, problem1 combine dataset, message = FALSE}
data_raintot = bind_rows(data_rain_2017, data_rain_2018) %>% 
    mutate(month = month.name[month]) %>% 
    select(year, month, total)

```

In trash_data, there are `r nrow(trash_data)` rows and `r ncol(trash_data)`columns with variables related to trash like plastic bottles. 
In precipitation data `data_rain`, there are `r nrow(data_raintot)` rows and `r ncol(data_raintot)` columns. Vraibles are month and the precipitation 
amount as "totoal" median are`r median(pull(trash_data%>% filter(year == 2017), sports_balls))`
    
   
# problem 2
## 2.1 First, clean the data in pols-month.csv. break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r, problem2 month-pol, message = FALSE}
data_pol =
  read_csv(file= "data/pols-month.csv")%>%
  janitor::clean_names()%>%
  separate(mon, into = c("year","month","day"), sep = "-", convert = TRUE) %>%
  mutate(month = month.abb[month]) %>%
   mutate(president = ifelse(prez_gop > 0, "gop","dem")) %>%
   select(-day, -prez_gop, - prez_dem)

```

## 2.2 Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.
```{r,problem2 snp, message = FALSE}
data_snp=read_csv(file = "data/snp.csv")%>%
  separate("date", into = c("day","month","year"), sep = "/", convert = TRUE)%>%
   arrange(year, month)%>%
  mutate(month = month.abb[month])%>%
  select(year, month, -day, everything())

```

## 2.3 import and tidy unemployment
```{r, problem2 unemployment, message = FALSE}
data_unemploy = read_csv(file="data/unemployment.csv")%>%
 pivot_longer(
             Jan:Dec,
             names_to = "month",
             values_to = "employment_value"
) %>% select(year = Year, everything())

 
  
```
## 2.4 Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r, problem2 combine dataset, message = FALSE}
data_combine= left_join(data_pol, data_snp, by = c("month","year"))
data_tot=left_join(data_combine,data_unemploy, by = c("month", "year"))
```
* For dataset data_pol, there are`r nrow(data_pol)` rows and `r ncol(data_pol)` columns. 
** Key vairables are time variables including `month` and `year`
** gov_gop: the number of republican governors on the associated date
** sen_gop: the number of republican senators on the associated date
** rep_gop: the number of republican representatives on the associated date
** gov_dem: the number of democratic governors on the associated date
** sen_dem: the number of democratic senators on the associated date
** rep_dem: the number of democratic representatives on the associated date

* For dataset data_snp, there are `r nrow(data_snp)` rows and `r ncol(data_snp)` columns. 
** Key variables are :
** date: the date of the observation
** close: the closing values of the S&P stock index on the associated date

* For dataset data_unemploy, there are `r nrow(data_unemploy)` and `r ncol(data_unemploy)`.
** Key variables are: unemployment_value: the percentage of unemployment of the associated month and year.

For the resulting dataset, there are `r nrow(data_tot)` rows and `r ncol(data_tot)` columns. the range of year is from `r min(pull(data_tot, year))` to `r max(pull(data_tot, year))`
key variables are 
** gov_gop: the number of republican governors on the associated date
** sen_gop: the number of republican senators on the associated date
** rep_gop: the number of republican representatives on the associated date
** gov_dem: the number of democratic governors on the associated date
** sen_dem: the number of democratic senators on the associated date
** rep_dem: the number of democratic representatives on the associated date
** date: the date of the observation
** close: the closing values of the S&P stock index on the associated date
** unemployment_value: the percentage of unemployment of the associated month an
which is a combine by the above three dataset by month and year.

# Problem 3
## 3.1 read and clean the dataset
```{r, problem3 baby name, message = FALSE}
data_baby=read_csv(file="data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP"= "BLACK NON HISPANIC","WHITE NON HISP" = "WHITE NON HISPANIC"))%>%
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>%
  distinct()
  
```

## 3.2 creat a reader friendly table
```{r, problem3 table, message = FALSE}
data_baby%>% filter(childs_first_name == "olivia",gender == "FEMALE") %>% select(childs_first_name, year_of_birth,ethnicity, rank) %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank") %>% knitr::kable()
## for male popular name
data_baby%>% filter(gender == "MALE",rank == 1) %>% select(childs_first_name, year_of_birth,ethnicity) %>% pivot_wider(names_from = "year_of_birth",
              values_from = "childs_first_name") %>% knitr::kable()
```

## 3.3 Creat a scatterplot
```{r, problem3 scatterplot, message = FALSE}
library(ggridges)
plot_df = filter(data_baby, ethnicity == "WHITE NON HISPANIC", gender == "MALE", year_of_birth == 2016)
ggplot(plot_df, aes(x = rank, y = count)) + geom_point() +
  labs(
    title = "The first name of white non-hispanic male children born in 2016",
    x = "The rank in popularity of that name",
    y = "The number of children with a name (count)")
```

